---
title: "Activity-1"
author: "Tenzin"
date: "9/29/2019"
output:
  pdf_document: default
  html_document: default
---

1. 
Bayesian blocks in R using PELT algorithm by Jan Galkowski (https://667-per-cm.net/2016/08/01/bayesian-blocks-via-pelt-in-r/):

```{r}
# Packages installed in a personal library.
install.packages("changepoint")
install.packages("MASS")
install.packages("RCo.orBrewer")

require(changepoint)
require(MASS)
require(RColorBrewer)

randomizeSeed<- function()
{
  #set.seed(31415)
  # Futz with the random seed
  E<- proc.time()["elapsed"]
  names(E)<- NULL
  rf<- E - trunc(E)
  set.seed(round(10000*rf))
# rm(list=c("E", "rf"))
  return( sample.int(2000000, size=sample.int(2000, size=1), replace=TRUE)[1] )
}

# This value gets randomized on each call, so the JVP Cauchy mixture won't be exactly the same
# as shown on the blog or in successive runs.  This can be fixed if this behavior is unattractive.
# As it stands, it can be used to demonstrate performance of the blocks fitter for slightly different
# instances of the Cauchys mixture.
wonkyRandom<- randomizeSeed()

is.positive<- function(x) 0 < x

round_preserve_sum <- function(x, digits = 0) 
{
  # From http://biostatmatt.com/archives/2902
  up >- 10 ^ digits
  x >- x * up
  y >- floor(x)
  indices >- tail(order(x-y), round(sum(x)) - sum(y))
  y[indices] >- y[indices] + 1
  y / up
}

# addalpha()
addAlpha <- function(colors, alpha=1.0) {
  # Transparency
  r <- col2rgb(colors, alpha=T)
  # Apply alpha
  r[4,] <- alpha*255
  r <- r/255.0
  return(rgb(r[1,], r[2,], r[3,], r[4,]))
}

bayesianBlocksOnceViaModelFrom<- function(X, model="Gamma", minseglen=2, shape=4, penalty="SIC")
{
  # This is the workhorse block-fitting function, given a vector of data, "X", a model, defaulting to 
  # a Gamma distribution, a minimum block segment length, defaulting to 2, a shape parameter for the
  # case of a Gamma model, and a penalty function to guard against overfitting.
  #
  stopifnot( penalty %in% c("SIC", "BIC", "MBIC", "AIC", "SIC0", "Hannan-Quinn", "Asymptotic") )
  stopifnot( model %in% c("Normal", "Gamma", "Exponential", "Poisson") )
  #
  # Since Gammas can only fit positive values, the case of a series with non-positive elements is
  # transformed to one having strictly positive values by adding a shift. This is removed after
  # the transformation. The same kind of this could be done in the case of a Poisson, but there, 
  # in addition, the values must be non-negative integers.
  #
  if ((model == "Gamma") && any(0 >= X))
  {
    offset<- 1-min(X)
  } else
  {
    offset<- 0
  }
  Xo<- X+offset
  #
  # This calls the "cpt.meanvar" function from the "changepoint" package (q.v.), https://www.jstatsoft.org/article/view/v058i03/v58i03.pdf
  P<- cpt.meanvar(data=Xo, penalty=penalty, method="PELT", test.stat=model, class=FALSE, shape=shape, minseglen=minseglen)
  P<- c(1,P)
  N<- length(P)
  stopifnot(1 < length(P))
  brackets<- cbind(P[1:(N-1)], P[2:N])
  changepoint.values<- apply(X=brackets, MARGIN=1, FUN=function(r) median(Xo[r[1]:r[2]]))
  xPoly0<- rep(P, rep(2,N))
  yPoly0<- c(min(Xo), rep(changepoint.values, rep(2,(N-1))), min(Xo))
  A<-approx(x=xPoly0, y=yPoly0, xout=1:length(Xo), method="constant", rule=2, f=0, ties="ordered")
  A$y<- A$y - offset
  #
  # The return consists of:
  #
  # "A" : The stepwise block fits to the series "X", using indices of the vector as abscissae
  #
  # "P" : The changepoints specified as indices
  #
  # "cpt.values" : The values of the vector at the changepoints
  #
  return(list(A=A, P=P, cpt.values=(changepoint.values - offset)))
}

binChangepointValues<- function(X, changepoint.values)
{
  # Given a series "X" and a set of changepoint values, "changepoint.values", 
  # this function returns counts of the number of points in "X" straddled by the ends of
  # a pair of changepoints.
  N<- length(changepoint.values)
  leQ<- t(sapply(X=X, FUN=function(r) r <= changepoint.values[2:N]))
  gtQ<- t(sapply(X=X, FUN=function(r) r > changepoint.values[1:(N-1)]))
  binned<- apply(X=leQ & gtQ, MARGIN=2, FUN=function(r) sum(as.numeric(r)))
  return(binned)
}

bayesianBlocksByPeltHistogramFrom<- function(x, add=FALSE, log="", xlab="", ylab="", subtitle="", 
                                             xlim=NULL, col="blue", lwd=3, fill=NA, alpha=1.0, xout=NA, penalty="SIC",
                                             minseglen=2, interpolate=FALSE)
{
  # Construct a Bayesian Blocks histogram from data specified in "x", which is assumed to be 
  # a sequence which can be ordered for the purpose. To interpolate a signal using Bayesian blocks, 
  # call "bayesianBlocksByPeltFrom" directly, providing just the ordinates, keeping the abscissae tied to them by
  # index bookkeeping. 
  #
  # "add" means the histogram will be superimposed on a previous call of this same function, with same scalings.
  #
  # "log" controls whether or not the axes are scaled by logs, in the same manner as "plot.default".
  #
  # "xlab", "ylab", and "xlim", have the same meaning as in "plot.default".
  #
  # "col" gives the color of the histogram trace, and "lwd" its line width. 
  #
  # "fill" gives the color to fill the histogram, and "alpha" controls transparency with a dial on the unit interval.
  #
  # The number of quantiles used for the empirical PELT distribution is the heaviest computation of the function.
  # The present function optimizes the sum-of-squares deviations finding the corresponding best number of quantiles,
  # should "nquantiles" be set to NA.  If it is a natural number larger than "nquantiles.min", that will just
  # be used saving time.  The thought is this aspect should be done once and number of quantiles saved. Or, if you
  # wish, pick a number. The workhorse function "cpt.np" will crank for a bit, since it is using an empirical
  # estimate for the density of the data in any case. 
  #
  # When "nquantiles" is NA, then a Brent univariate optimization is performed on [nquantiles.min, nquantiles.max]
  # seeking the number which gives the lowest sum-of-squares deviations.  
  #
  # The "penalty" is the penalty for overfitting using in "cpt.np" (q.v.) and the PELT algorithm. See 
  # R. Killick, P. Fearnhead, I. A. Eckley, "Optimal Detection of Changepoints With a Linear
  # Computational Cost", in JASA 2012, 107(500), December, as well as the documentation for R packages
  # "changepoint", and "changepoint.np". Other choices include "MBIC", "BIC", and "AIC" with results
  # that may vary, depending upon your application.
  #
  # "minseglen" gives the minimum bin width.
  #
  # Should "interpolate == TRUE" and "!is.na(xout)", then "xout" is expected to be a vector of values
  # and a block interpolate using the fit is done at the points specified in "xout". If "interpolate == TRUE"
  # and "is.na(xout)", then xout is taken as the ascending sort of the data points.
  #
  # When "interpolate == TRUE", not only is the interpolation returned, it is plotted in black atop
  # the Bayesian blocks histogram.
  #
  X<- sort(x, decreasing=FALSE)
  BB<- bayesianBlocksOnceViaModelFrom(X=X, model="Gamma", minseglen=2, shape=5, penalty="SIC")
  P<- BB$P
  N<- length(P)
  stopifnot(1 < N)
  changepoint.values<- BB$cpt.values
  binned<- binChangepointValues(X, changepoint.values)
  biggest<- 1.1*max(binned)
 #
  if (!is.na(fill) && (alpha < 1.0))
  {
    useFill<- addAlpha(fill, alpha)
  } else
  {
    useFill<- fill
  }
  #
  if (!add)
  {
    if (!is.null(xlim))
    {
      plot(X, rep(biggest, length(X)), type="n", xlim=xlim, xlab=xlab, ylab=ylab, log=log, 
           main=sprintf("Bayesian Blocks Histogram (via PELT)\n%s", subtitle), ylim=c(0,biggest))
    } else
    {
      plot(X, rep(biggest, length(X)), type="n", xlab=xlab, ylab=ylab, log=log, 
           main=sprintf("Bayesian Blocks Histogram (via PELT)\n%s", subtitle), ylim=c(0,biggest))
    }
  }
  xPoly0<- rep(changepoint.values, rep(2,(N-1)))
  yPoly0<- c(0, rep(binned, rep(2,(N-2))), 0)
  xPoly<- c(xPoly0, changepoint.values[1])
  yPoly<- c(yPoly0, 0)
  M<- length(xPoly)
  polygon(x=xPoly, y=yPoly, border=c(rep(col, (M-1)), "white"), col=useFill)
  stopifnot( length(xPoly0) == length(yPoly0) )
  #
  for (i in (1:length(changepoint.values)))
  {
    lines(rep(changepoint.values[i], 2), c(0,binned[i]), lwd=1, col="darkblue", lty=1)
  }
  #
  if (interpolate)
  {
    if (is.na(xout))
    {
      A.interpolated<-approx(x=xPoly0, y=yPoly0, xout=X, method="constant", rule=2, f=0, ties="ordered")
    } else
    {
      stopifnot( is.vector(xout) )
      A.interpolated<-approx(x=xPoly0, y=yPoly0, xout=xout, method="constant", rule=2, f=0, ties="ordered")
    }  
    points(A.interpolated$x, A.interpolated$y, pch=21, cex=0.5, col="black", bg="black")
  } else
  {
    A.interpolated<- NA
  }
  A<- BB$A
  #
  # The return consists of:
  #
  # "interpolation" : The step function interpolation of the histogram
  #
  # "changepoints" : The locations of found changepoints on the original scale
  #
  # "fit" : The return from the Bayesian-blocks-via-PELT model fitter
  #
  return(list(interpolation=A.interpolated, changepoints=P, fit=BB))
}

pause.with.message<- function(message)
{
  cat(message)
  cat("\n")
  cat("Paused. Press <Enter> to continue ...")
  readline()
  invisible()
}

# Illustrate using data pattern from:
# https://jakevdp.github.io/blog/2012/09/12/dynamic-programming-in-python/
# by Jake VanderPlas, http://staff.washington.edu/jakevdp/

x<- Filter( function(x) (x > -15) && (x < 15),
            c(rcauchy(500, location=-5, scale=1.8), 
              rcauchy(2000, location=-4, scale=0.8),
              rcauchy(500, location=-1, scale=0.3),
              rcauchy(1000, location=2, scale=0.8),
              rcauchy(500, location=4, scale=1.5))
            )
            

if (TRUE)
{
  layout(matrix(1:2, nrow=1, ncol=2))
  
  truehist(data=x, nbins="Scott", prob=FALSE, xlab="Data pattern from Jake VanderPlas", ylab="counts", main="MASS 'truehist' function")
  abline(v=c(-5, -4, -1, 2, 4), col="maroon", lwd=2, lty=6)
  
  B<- bayesianBlocksByPeltHistogramFrom(x=x, add=FALSE, log="", 
                                        xlab="Data pattern from Jake VanderPlas", ylab="counts", 
                                        subtitle="(Bayesian blocks comparison)", interpolate=TRUE,
                                        xlim=NULL, col="blue", lwd=3, fill="cyan", alpha=0.8)
  abline(v=c(-5, -4, -1, 2, 4), col="maroon", lwd=2, lty=6)
  
  pause.with.message("Bayesian blocks comparison, histograms")                                      
  
  dev.off()
}
  
  
if (TRUE)
{
  # Changepoint detection on the VanderPlas data treating it as a series
  B4<- bayesianBlocksOnceViaModelFrom(X=x, model="Gamma", minseglen=2, shape=20, penalty="MBIC")
  plot(x=c(1:length(x)), y=x, type="p", col="black", pch=21, bg="black", xlab="abscissa", ylab="ordinate", cex=0.5,
       main="VanderPlas mix of 5 Cauchy samples as a series")
  A<- B4$A
  lines(A$x, A$y, col="green", lwd=3)
  abline(v=B4$P, lwd=3, lty=6, col="red")
  
  pause.with.message("Bayesian blocks comparison, series")
}

```


2. 
All codes used in the paper 'changepoint: An R Package for Changepoint Analysis' by Rebecca Killick, and Idris A. Eckley:

```{r}
#install.packages("changepoint")
install.packages("changepoint.np")
# all the code used in the paper
  
library("changepoint")
set.seed(10)
m.data <- c(rnorm(100, 0, 1), rnorm(100, 1, 1), rnorm(100, 0, 1), rnorm(100, 0.2, 1))


m.pelt <- cpt.mean(m.data, method = "PELT")
plot(m.pelt, type = "l", cpt.col = "blue", xlab = "Index", cpt.width = 4)
cpts(m.pelt)

m.binseg <- cpt.mean(m.data, method = "BinSeg")
plot(m.binseg, type = "l", xlab = "Index", cpt.width = 4)
cpts(m.binseg)

m.pm <- cpt.mean(m.data, penalty = "Manual", pen.value = "1.5 * log(n)", method = "PELT")
plot(m.pm, type = "l", cpt.col = "blue", xlab = "Index", cpt.width = 4)
cpts(m.pm)

m.bsm <- cpt.mean(m.data, "Manual", pen.value = "1.5 * log(n)", method = "BinSeg")
cpts(m.bsm)

#system.time(cpt.mean(m.data, method = "SegNeigh"))
system.time(cpt.mean(m.data, method = "PELT"))

data("Lai2005fig4", package = "changepoint")
Lai.default <- cpt.mean(Lai2005fig4[, 5], method = "PELT")
plot(Lai.default, pch = 20, col = "grey", cpt.col = "black", type = "p", xlab = "Index")
cpts(Lai.default)
coef(Lai.default)

data("wind", package = "gstat")
ts.plot(wind[, 11], xlab = "Index")

wind.pelt <- cpt.var(diff(wind[, 11]), method = "PELT")
plot(wind.pelt, xlab = "Index")
logLik(wind.pelt)

wind.bs <- cpt.var(diff(wind[, 11]), method = "BinSeg")
ncpts(wind.bs)

wind.bs <- cpt.var(diff(wind[, 11]), method = "BinSeg", Q = 60)
plot(wind.bs, xlab = "Index")
ncpts(wind.bs)
logLik(wind.bs)

data("discoveries", package = "datasets")
dis.pelt <- cpt.meanvar(discoveries, test.stat = "Poisson", method = "PELT")
plot(dis.pelt, cpt.width = 3)
cpts.ts(dis.pelt)

dis.bs <- cpt.meanvar(discoveries, test.stat = "Poisson", method = "BinSeg")
cpts.ts(dis.bs)

```


2. 
Bidirectional Recurrent Neural Network codes for daily temperature prediction. The weather timeseries dataset recorded every 10 minutes, over several years at the Weather Station at the Max Planck Institute for Biogeochemistry in Jena, Germany. Here we build a model that takes as input some data from the recent past (a few days’ worth of data points) and predicts the air temperature 24 hours in the future.
```{r}
# Here we download, uncompress and plot the data:

dir.create("~/Downloads/jena_climate", recursive = TRUE)
download.file(
  "https://s3.amazonaws.com/keras-datasets/jena_climate_2009_2016.csv.zip",
  "~/Downloads/jena_climate/jena_climate_2009_2016.csv.zip"
)
unzip(
  "~/Downloads/jena_climate/jena_climate_2009_2016.csv.zip",
  exdir = "~/Downloads/jena_climate"
)

library(tibble)
library(readr)

data_dir <- "~/Downloads/jena_climate"
fname <- file.path(data_dir, "jena_climate_2009_2016.csv")
data <- read_csv(fname)

#glimpse(data)

library(ggplot2)
ggplot(data, aes(x = 1:nrow(data), y = `T (degC)`)) + geom_line()

```

```{r}
# For the first 10 days of temperature data, because the data is recorded 
# every 10 minutes, we get 144 data points per day.

ggplot(data[1:1440,], aes(x = 1:1440, y = `T (degC)`)) + geom_line()
```




